Computational methods in landscape ecology

### 2024-09-20 Manuscript.qmd
Fix format issues

papersize: a4paper
mainfont: Garamond
fontsize: 11pt
geometry: margin=25mm
indent: 2m
linestretch: 1.5
number-sections: false

bibliography: CLER.bib
csl: springer-basic-author-date.csl
---

@@

@@

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-spatial-patterns.qmd >}}
{{< include Sections/03-connectivity.qmd >}}
{{< include Sections/04-sampling.qmd >}}
{{< include Sections/05-simulations.qmd >}}
{{< include Sections/06-genetics.qmd >}}
{{< include Sections/07-spatial-planning.qmd >}}
{{< include Sections/08-discussion.qmd >}}

# Acknowledgments

## Funding

## Author contributions

All authors contributed particular sections to the manuscript coordinated by MHKH.
Correspondence to [Maximilian H.K. Hesselbarth](mailto:mhk.hesselbarth@gmail.com)

## Conflict of Interest

## Human and Animal Rights and Informed Consent

>>>>

papersize: a4paper
fontsize: 11pt
geometry: margin=25mm
indent: 2m
linestretch: 1.5
number-sections: false

bibliography: CLER.bib
csl: springer-basic-brackets-no-et-al.csl
---

@@

\noindent \textbf{Acknowledgments}

\noindent \textbf{Funding}

\noindent \textbf{Author contributions}

All authors contributed particular sections to the manuscript coordinated by MHKH.

Correspondence to [Maximilian H.K. Hesselbarth](mailto:mhk.hesselbarth@gmail.com)

\noindent \textbf{Conflict of Interest}

\noindent \textbf{Human and Animal Rights and Informed Consent}

@@

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-spatial-patterns.qmd >}}
{{< include Sections/03-connectivity.qmd >}}
{{< include Sections/04-sampling.qmd >}}
{{< include Sections/05-simulations.qmd >}}
{{< include Sections/06-genetics.qmd >}}
{{< include Sections/07-spatial-planning.qmd >}}
{{< include Sections/08-conclusion.qmd >}}

# Key references

**Frazier (2019) Emerging trajectories for spatial pattern analysis in landscape ecology. Landscape Ecology 34:2073–2082.** The paper describes how interdisciplinary perspectives have contributed to spatial pattern analysis and introduces new innovative developments.

**Unnithan Kumar et al. (2022) Moving beyond landscape resistance: Considerations for the future of connectivity modelling and conservation science. Landscape Ecology 37:2465–2480.** The paper provides a historical overview of connectivity modelling with a focus on resistance surfaces and discusses current missing aspects and how to address them.

**Zuckerberg et al. (2020) A review of overlapping landscapes: Pseudoreplication or a red herring in landscape ecology? Current Landscape Ecology Reports 5:140–148.** The paper discusses theoretical and practical implications over overlapping landscapes in the context of landscape buffers of varying sizes surrounding sampling sites.

**Synes et al. (2016) Emerging opportunities for landscape ecological modelling. Current Landscape Ecology Reports 1:146–167.** The paper identifies existing gaps in landscape ecological modelling while highlighting potential emerging opportunities.

**Manel and Holderegger (2013) Ten years of landscape genetics. Trends in Ecology & Evolution 28:614–621.** Ten years after their first paper introducing landscape genetics, this paper reviews the state-of-the-art and describes main topics that have contributed to the progress of landscape genetics.

**Jung et al. (2024) An assessment of the state of conservation planning in europe. Philosophical Transactions of the Royal Society B 379:20230015.** The paper assesses where and how spatial conservation planning have been applied across Europe.

### 2024-09-20 Sections/06-genetics.qmd
Fix format issues

Advances in molecular methodologies and synergistic developments of bioinformatic and computational approaches to analyze large-scale genomic data have recently enabled the integration of genome-wide data with spatial ecological data.<!-- MH: Ref missing -->.
This shift towards landscape genomics [@Cushman2018EditorialLeast; @Storfer2018NavigatingInterface] comes with novel utilities through which researchers can apply genomic data across spatio-temporal scales, including the integration of historic or ancient DNA [@Fenderson2020SpatiotemporalLandscape], or simulated present-day and future population genomic data [@House2018EvaluatingMethods; @Chevy2024PopulationGenetics].
Analytical frameworks such as the *FOLDS* model [gene flow, genetic offsets, genetic load, dispersal and SDMs\; @Aguirre-Liguori2021EvolutionaryGenomics], constrained coordination where covarying sets of genotypes are correlated with multivariate environments through redundancy analysis [RDA\; @Capblancq2021RedundancyAnalysis], and spatial simulations provide promising potential for future landscape genomic studies.
Such studies may explore interactions between complex evolutionary processes, including demography, multidimensional (e.g., strength and direction) gene flow and migration across spatio-temporal scales, genetic differentiation across landscapes, genomic load and adaptive potential, and the interpretation of these evolutionary processes in context of landscape ecology [@Balkenhol2017LandscapeGenomics; @House2018EvaluatingMethods; @Fenderson2020SpatiotemporalLandscape; @Marcus2021FastFlexible; @Chevy2024PopulationGenetics].

>>>>

Advances in molecular methodologies and synergistic developments of bioinformatic and computational approaches to analyze large-scale genomic data have recently enabled the integration of genome-wide data with spatial ecological data.
This shift towards landscape genomics [@Cushman2018EditorialLeast; @Storfer2018NavigatingInterface] comes with novel utilities through which researchers can apply genomic data across spatio-temporal scales, including the integration of historic or ancient DNA [@Fenderson2020SpatiotemporalLandscape], or simulated present-day and future population genomic data [@House2018EvaluatingMethods; @Chevy2024PopulationGenetics].
Analytical frameworks such as the *FOLDS* model [gene flow, genetic offsets, genetic load, dispersal and SDMs\; @Aguirre-Liguori2021EvolutionaryGenomics], constrained coordination where covarying sets of genotypes are correlated with multivariate environments through redundancy analysis [RDA\; @Capblancq2021RedundancyAnalysis], and spatial simulations provide promising potential for future landscape genomic studies.
Such studies may explore interactions between complex evolutionary processes, including demography, multidimensional (e.g., strength and direction) gene flow and migration across spatio-temporal scales, genetic differentiation across landscapes, genomic load and adaptive potential, and the interpretation of these evolutionary processes in context of landscape ecology [@Balkenhol2017LandscapeGenomics; @House2018EvaluatingMethods; @Fenderson2020SpatiotemporalLandscape; @Marcus2021FastFlexible; @Chevy2024PopulationGenetics].
